Sufficient Conditions for Reachability in Automata Networks with Priorities

### 2014-12-18 parts/sa.tex
revise GLC + explanations

The graph of local causality $\cwB = (\V, \E)$ defined in \pref{def:glc} is a graph where $\V \subseteq \Proc \cup \Obj \cup \Sol$ and $\E \subseteq \V \times \V$.
A node in $\PHproc$ is a required process,
a node in $\Obj$ is an objective to reach a given process
and a node in $\Sol$ is a set of processes required for the solving.
An objective $P \in \Obj$ is solvable if the abstractions of bounce sequences $\aBS(P) \in \Sol$ (\pref{def:aBS}) can be reached (\pref{eq:ESol1}), thus leading to several required processes (\pref{eq:ESol2}).
If $a \in \components$, the reachability of one of its process $a_i$ is approximated by the ability to solve all objectives $\PHobjp{a}{j}{i} \in \Obj$ for all $a_j$ in the initial context (\pref{eq:EReq});
if $a \in \cs$, the reachability of $a_i$ is simply solved by the set of processes $\csState(a_i)$ (\pref{def:csState}) that it represents (\pref{eq:EPrio}).
The solving of an objective $P$ may require a process of $\PHsort(P)$, \ie $\gCont(\PHsort(P), P) \neq \emptyset$ (\pref{def:maxCont}); in this case, $P$ is re-targeted (\pref{eq:ECont}).
\pref{eq:Vw}, \eqref{eq:Vproc} and~\eqref{eq:VE} ensure that all required nodes are in $\cwV$.
Finally, as the active process of every sort may evolve, $\cwB$ is obtained by iteratively saturating with every process it contains,
\ie by overriding its initial context $\ctx$ by $\allprocs(\V, \E)$, defined by:
\[\allprocs(\V, \E) = (V \cap \Proc) \cup \{ \PHtarget(P), \PHbounce(P) \mid P \in \V \cap \Obj \} \enspace.\]

\begin{definition}
\label{def:glc}
  The graph of local causality $\cwB \DEF (\Bv, \Be)$ is defined as: $\cwB \DEF \lfp{\aB^\w_\ctx}{\myB}{\aB^\w_{\ctx \Cap \allprocs(\myB)}}$,
  where $\myB \DEF (\cwV, \cwE)$ is the smallest graph with
  $\cwV \subseteq \Proc \cup \Obj \cup \Sol$ and $\cwE \subseteq \cwV \times \cwV$
  so that:
  \begin{align}
    \w &\subseteq \cwV \label{eq:Vw} \\
    P \in \cwVObj &\Rightarrow \PHbounce(P) \in \cwV \label{eq:Vproc} \\
    (x, y) \in \cwE &\Rightarrow y \in \cwV \label{eq:VE} \\
    P \in \cwVObj \wedge ps \in \modMF{\aBS}(P) &\Rightarrow (P, ps) \in \cwE \label{eq:ESol1} \\
    ps \in \cwVSol \wedge a_i \in ps &\Rightarrow (ps, a_i) \in \cwE \label{eq:ESol2} \\
    a \in \components \wedge a_i \in \cwVProc \wedge a_j \in \ctx &\Rightarrow (a_i, \PHobjp{a}{j}{i}) \in \cwE \label{eq:EReq} \\
    a \in \cs \wedge a_i \in \cwVProc \wedge ps \in \csState(a_i) &\Rightarrow (a_i, ps) \in \cwE \label{eq:EPrio} \\
    P \in \cwVObj \wedge q \in \gCont_\ctx(\PHsort(P), P) &\Rightarrow (P, \PHobj{q}{\PHbounce(P)}) \in \cwE \label{eq:ECont} \!
  \end{align}
\end{definition}

In the graph of local causality, an edge $(p, ps) \in \Proc \times \Sol$
is said to be coherent (\pref{def:coherent}) if none of the processes in $ps$ conflict with the children processes of $ps$.
Then, \pref{th:approxinf} gives a sufficient condition for the concretisation of a sequence of objectives in a given context,
which is derived immediately from the graph of local causality.
A proof of this theorem is given in \pref{suppl:demoapproxinf}.
\begin{definition}[Coherent edge]
\label{def:coherent}
  An edge $(x, y) \in \cwE$ is said to be coherent if and only if:
  $(x, y) \in \Be \cap (\Proc \times \Sol) \Rightarrow y$ has no children process $a_j \in \Bv \cap \Proc$ so that $\exists a_i \in y$, $a_i \neq a_j$.
\end{definition}

@@

\begin{figure}[tp]
  \centering
  \begin{tikzpicture}[aS]
    \node[Aproc] (c1) {$c_1$};
    \node[Aobj,below of=c1] (c01) {$\PHobj{c_0}{c_1}$};
    \node[Asol,below of=c01] (c01s) {};

@@

    \path
    (c1) edge (c01)
    (c01) edge (c01s)
    (c01s) edge (ab11)
    (ab11) edge[aSPrio] (ab11s)
    (ab11s) edge (a1) edge (b1)

@@

    (b0) edge (b10) edge (b00)
    (b10) edge (b10s)
    (b00) edge (b00s)
    (b1) edge (b01) edge (b11)
    (b01) edge (b01s)
    (b01s) edge (a0)
    (b11) edge (b11s)
    ;
    \end{tikzpicture}
  \caption{
  \label{fig:sa-livelock}
    The graph of local causality of the PH model in \pref{fig:ph-livelock}
    for the objective $\w = \PHobj{c_0}{c_1}$ and the initial context $\ctx = \PHstate{a_1, b_0, c_0, ab_{10}}$.
    Rectangular nodes containing a single process are elements in $\Proc$,
    borderless nodes containing a couple of processes are elements in $\Obj$
    and circle nodes are elements in $\Sol$.
    \pref{th:approxinf} is inconclusive on this example as edge $(ab_{11}, \{ a_1, b_1 \}) \in \Proc \times \Sol$ (here represented with a double line) is not coherent (\pref{def:coherent}).
    Indeed, $a_0 \in \Proc$ is a child of $\{ a_1, b_1 \}$, but $a_0 \neq a_1$ (and the same also stands for $b_0$).
  }
\end{figure}
\end{example}

>>>>

\modMF{
A graph of local causality $\aB = (\cwV, \cwE)$
is a graph where $\cwV \subseteq \Proc \cup \Obj \cup \Sol$ is the set of vertices
and $\cwE \subseteq \cwV \times \cwV$ is the set of oriented edges.
Such a graph aims a describing the requirements of a given reachability
problem, that is, the reachability of $\w$ from the initial context $\ctx$.
Thus, a node in $\PHproc$ represents a process required to play an action,
a node in $\Obj$ is an objective to reach a given process
and a node in $\Sol$ is a set of hitter processes required for the solving
of an objective.
}

\modMF{
The edges of such a graph allow to draw links between these requirements.
Thus, an objective $P \in \Obj$ is solvable if
at least one related bounce sequence in $\BS(P)$ can be played;
therefore, we can consider all minimal sets of hitters of these bounce sequences
(that is, the sets of $\aBS(P) \in \Sol$, see \pref{def:aBS})
as requirements to solve this objective (\pref{eq:ESol1}).
Such a set can then naturally be split into as many process nodes,
making each of them a new requirement (\pref{eq:ESol2}).
If $a \in \components$, the reachability of a required process $a_i$
is approximated by the ability to solve all objectives of the form
$\PHobjp{a}{j}{i} \in \Obj$ for all $a_j$ in the initial context (\pref{eq:EReq});
if $a \in \cs$, the reachability of $a_i$ is simply solved by the set of processes $\csState(a_i)$ (see \pref{def:csState}) that it represents (\pref{eq:EPrio}).
Furthermore, we note that the solving of an objective $P$ may require
a process of $\PHsort(P)$,
\ie $\gCont(\PHsort(P), P) \neq \emptyset$ (\pref{def:maxCont});
in this case, $P$ is re-targeted (\pref{eq:ECont}).
Finally, \pref{eq:Vw} forces the graph to contain all the initial objectives,
and \pref{eq:VE} ensures the consistency of $\cwV$.
}

\modMF{
However, this graph $\aB$ alone is not sufficient to check the reachability of
an objective sequence $\w$ from an initial context $\ctx$;
indeed, the reachability of each intermediate required process
is only checked starting from the context $\ctx$,
although this context may be modified by other objectives.
Therefore, such a reachability requires the building of the
\emph{saturated graph of local causality} (\pref{def:glc})
which is obtained by iteratively saturating the initial context with all
processes referenced in the graph,
until reaching a fixed point.
Such a fixed point always exists because the size of the context
is limited by the size of the initial PH model.
The set of processes contained in a graph of local causality
is given by:
\begin{align*}
  \allprocs(\cwV, \cwE) = (\cwV \cap \Proc) &\cup \{ \PHtarget(P) \mid P \in \cwV \cap \Obj \}\\
  & \cup \{ P \neq \w \Rightarrow \PHbounce(P) \mid P \in \cwV \cap \Obj \}
\end{align*}
where, if the objective sequence $\w$ consists of only one element,
then we can omit the hitter of this objective
if it does not appear elsewhere in the graph.
}

\begin{definition}
\label{def:glc}
  The graph of local causality $\cwB \DEF (\Bv, \Be)$ is defined as: $\cwB \DEF \lfp{\aB^\w_\ctx}{\myB}{\aB^\w_{\ctx \Cap \allprocs(\myB)}}$,
  where $\myB \DEF (\cwV, \cwE)$ is the smallest graph with
  $\cwV \subseteq \Proc \cup \Obj \cup \Sol$ and $\cwE \subseteq \cwV \times \cwV$
  so that:
  \begin{align}
    \w &\subseteq \cwV \label{eq:Vw} \\
%    P \in \cwVObj &\Rightarrow \PHbounce(P) \in \cwV \label{eq:Vproc} \\
    (x, y) \in \cwE &\Rightarrow y \in \cwV \label{eq:VE} \\
    P \in \cwVObj \wedge ps \in \modMF{\aBS}(P) &\Rightarrow (P, ps) \in \cwE \label{eq:ESol1} \\
    ps \in \cwVSol \wedge a_i \in ps &\Rightarrow (ps, a_i) \in \cwE \label{eq:ESol2} \\
    a \in \components \wedge a_i \in \cwVProc \wedge a_j \in \ctx &\Rightarrow (a_i, \PHobjp{a}{j}{i}) \in \cwE \label{eq:EReq} \\
    a \in \cs \wedge a_i \in \cwVProc \wedge ps \in \csState(a_i) &\Rightarrow (a_i, ps) \in \cwE \label{eq:EPrio} \\
    P \in \cwVObj \wedge q \in \gCont_\ctx(\PHsort(P), P) &\Rightarrow (P, \PHobj{q}{\PHbounce(P)}) \in \cwE \label{eq:ECont} \!
  \end{align}
\end{definition}

\todo{“Saturated” partout}

In the saturated graph of local causality, an edge $(p, ps) \in \Proc \times \Sol$
is said to be coherent (\pref{def:coherent}) if none of the processes in $ps$ conflict with the children processes of $ps$.
Then, \pref{th:approxinf} gives a sufficient condition for the concretisation of a sequence of objectives in a given context,
which is derived immediately from the graph of local causality.
A proof of this theorem is given in \pref{suppl:demoapproxinf}.
\begin{definition}[Coherent edge]
\label{def:coherent}
  An edge $(x, y) \in \cwE$ is said to be coherent if and only if:
  $(x, y) \in \Be \cap (\Proc \times \Sol) \Rightarrow y$ has no children process $a_j \in \Bv \cap \Proc$ so that $\exists a_i \in y$, $a_i \neq a_j$.
\end{definition}

@@

\begin{figure}[tp]
  \centering
  \begin{tikzpicture}[aS]
%    \node[Aproc] (c1) {$c_1$};
%    \node[Aobj,below of=c1] (c01) {$\PHobj{c_0}{c_1}$};
    \node[Aobj] (c01) {$\PHobj{c_0}{c_1}$};
    \node[Asol,below of=c01] (c01s) {};

@@

    \path
%    (c1) edge (c01)
    (c01) edge (c01s)
    (c01s) edge (ab11)
    (ab11) edge[aSPrio] (ab11s)
    (ab11s) edge (a1) edge (b1)

@@

    (b0) edge (b10) edge (b00)
    (b10) edge (b10s)
    (b00) edge (b00s)
    (b1) edge (b01) edge (b11)
    (b01) edge (b01s)
    (b01s) edge (a0)
    (b11) edge (b11s)
    ;
    \end{tikzpicture}
  \caption{
  \label{fig:sa-livelock}
    The saturated graph of local causality of the PH model in \pref{fig:ph-livelock}
    for the objective $\w = \PHobj{c_0}{c_1}$ and the initial context $\ctx = \PHstate{a_1, b_0, c_0, ab_{10}}$.
    Rectangular nodes containing a single process are elements in $\Proc$,
    borderless nodes containing a couple of processes are elements in $\Obj$
    and circle nodes are elements in $\Sol$.
    \pref{th:approxinf} is inconclusive on this example as edge $(ab_{11}, \{ a_1, b_1 \}) \in \Proc \times \Sol$ (here represented with a double line) is not coherent (\pref{def:coherent}).
    Indeed, $a_0 \in \Proc$ is a child of $\{ a_1, b_1 \}$, but $a_0 \neq a_1$ (and the same also stands for $b_0$).
  }
\end{figure}
\end{example}
